---
title: "Assignment 3"
author: "Dylan Culp"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

My four datasets are from Chicago. The first two are from the CTA open data site (https://www.transitchicago.com/data/) and the second two are from the city of Chicago's Open Data portal (https://data.cityofchicago.org/)

  A. CTA bus stops (point)
  B. CTA "L" Stations (point)
  C. Community Areas (polygon)
  D. Central Business District (polygon)
  
  
I will bring each dataset into R: 
```{r}
bus_stops <- st_read("https://data.cityofchicago.org/download/84eu-buny/text%2Fxml", quiet=TRUE)
l_stops <- st_read("https://data.cityofchicago.org/download/4qtv-9w43/application%2Fxml", quiet=TRUE)
comm_areas <- st_read("https://data.cityofchicago.org/api/geospatial/cauq-8yn6?method=export&format=GeoJSON", quiet=TRUE)
cbd <- st_read("https://data.cityofchicago.org/api/geospatial/tksj-nvsw?method=export&format=GeoJSON", quiet=TRUE)
```
  
  
I will then transform the coordinates for each dataset to a Chicago PROJ4 string: 
```{r}
chicago_coordinates <- "+proj=tmerc +lat_0=36.66666666666666 +lon_0=-88.33333333333333 +k=0.9999749999999999 +x_0=300000 +y_0=0 +ellps=GRS80 +units=m +no_defs"

bus_stops <- bus_stops %>%
  st_transform(chicago_coordinates)

l_stops <- l_stops %>%
  st_transform(chicago_coordinates)

comm_areas <- comm_areas %>%
  st_transform(chicago_coordinates)

cbd <- cbd %>%
  st_transform(chicago_coordinates)
```

1. How many L stops are in each community area?

```{r}
comm_areas <- comm_areas %>%
  mutate(l_stops = lengths(st_covers(comm_areas, l_stops)))

ggplot(comm_areas)+
  geom_sf(color = NA,
  aes(fill = l_stops))
```

2. How many bus stops are in each community area?

```{r}
comm_areas <- comm_areas %>%
  mutate(bus_stops = lengths(st_covers(comm_areas, bus_stops)))

ggplot(comm_areas)+
  geom_sf(color = NA,
  aes(fill = bus_stops))
```

3. Average distance between an L stop and nearest bus stop?

```{r}
l_stops <- l_stops %>%
  mutate(l_dist = st_nn(l_stops, bus_stops,
                          returnDist = TRUE)$dist) %>%
    mutate(l_dist = as.numeric(l_dist))

avg_bus_dist <-mean(l_stops$l_dist)
avg_bus_dist
```
```{r}
ggplot(water) + 
  geom_sf(data = bus_stops, size = 0.1,
          aes(color=avg_bus_dist))+
  scale_color_viridis_c(name = "Distance between bus stops and L stops")

```




4. How many neighborhoods does the CBD overlap with?

```{r}
comm_areas <- comm_areas %>%
  mutate(cbd_overlap = lengths(st_overlaps(cbd, comm_areas))) %>%
  mutate(has_overlap = cbd_overlap > 0)

```

Chicago's CBD overlaps with 5 neighborhoods.


```{r}
ggplot()+
  geom_sf(data = comm_areas)
  geom_sf(data = comm_areas)
  
```

