---
title: "Assignment 2"
author: "Dylan Culp"
date: "9/22/2020"
output: html_document
---

Here are my libraries:

```{r}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(rgeos)
library(rnaturalearthdata)
```

I've chosen Africa as my area to study. I will import the world dataset, then filter out all non-African countries:

```{r}
world <- ne_countries(scale = 50, returnclass = "sf")
africa_countries <- world %>%
                    filter(continent=="Africa")

```

This first map is using an Albers Equal Ares Africa set of coordinates
```{r}
ggplot(africa_countries)+
       geom_sf()+
      theme_bw()
```
I will try two other coordinate systems: an equal area focusing on Brazil, and a mercator for India

```{r}
brazil_coords <- "+proj=aea +lat_1=10 +lat_2=-40 +lat_0=-25 +lon_0=-50 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

india_coords <- "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext  +no_defs"


africa_countries_2 <- africa_countries %>%
  st_transform(crs = brazil_coords)

africa_countries_3 <- africa_countries %>%
  st_transform(crs = india_coords)
```

Here are those maps:

```{r}
ggplot(africa_countries_2)+
       geom_sf()+
      theme_bw()

ggplot(africa_countries_3)+
       geom_sf()+
      theme_bw()
```



I will then use this dataset for my first map--a standard chloropleth map, illustrating the population estimates per country. I've chosen an Albers Equal Area Africa projection.

```{r}
africa_aea <- "+proj=aea +lat_1=36.5 +lat_2=29.071428571429 +lat_0=32.7857142857145 +lon_0=-14.111111111111 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

africa_pop_map <- ggplot(africa_countries)+
                  geom_sf(aes(fill = pop_est))
                  coord_sf(crs = africa_aea)

africa_pop_map                  
```

My second map will be a cartogram that retains the above color scheme for population, while adjusting each country's size according to GDP.

Note: I believe there is an error here--I think the distortions are due to the coordinates rathern than the actual 

```{r}
ggplot(africa_countries, aes(fill = pop_est)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Population by country in Africa",
    breaks = seq(0, 200000000, by = 20000000),
    labels = formatC(seq(0, 200000000, by = 20000000), 
                     big.mark = ",", format = "f", digits = 0)) +
  coord_sf(crs = africa_aea)+
  theme_map() +
  theme(legend.background = element_blank())
```

Lastly, I'd like to arrange the countries in order, from largest in population to smallest.I had trouble getting the labels in a way that has both readable text and is positioned correctly under the corresponding column.


```{r}
ggplot(africa_countries) +
  geom_col(aes(x=reorder(sovereignt, -pop_est), y= pop_est))+
          theme(axis.text.x = element_text(size=6, angle=315))
             
```

