---
title: "Assignment 4"
author: "Dylan Culp"
date: "10/6/2020"
output: html_document
---

```{r}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
```

I will use a dataset of recycling drop off sites in Chicago for this exercise. 

```{r}
recycling_sites <- st_read("https://data.cityofchicago.org/api/geospatial/27cy-fdic?method=export&format=KML")
recycling_sites <- recycling_sites[-1,]
```
Here is a map of Chicago's streets from OSM:

```{r}
chicago_street_features <- opq(bbox = 'Chicago IL USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

chicago_streets <- chicago_street_features$osm_lines

ggplot(chicago_streets)+
  geom_sf()+
  theme_map()
```
Here is the groundwork for building the isochrones: 
```{r}
path_otp <- otp_dl_jar("OTP")
```

```{r}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r}
otp_setup(otp = path_otp, dir = path_data, memory =1024)

```

```{r}
otpcon <- otp_connect()

```


My first map overlays 10 minute walk and 10 minute drive isochrones:
```{r}
iso_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = recycling_sites, 
                mode = "WALK", cutoffSec = 600) %>%
  mutate(mode = "walk")

iso_10min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = recycling_sites, 
                mode = "CAR", cutoffSec = 600) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_10min_drive, iso_10min_walk)

otp_stop()

```


Here is the map output:
```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = recycling_sites) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 10 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```
My next map will look at bike accessibility within 10 minutes:
```{r}
iso_10min_bike <- 
  otp_isochrone(otpcon = otpcon, fromPlace = recycling_sites, 
                mode = "BICYCLE", cutoffSec = 600) %>%
  mutate(mode = "bicycle")

right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_10min_bike) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = recycling_sites) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 10 minutes",
                       labels = c("By bicycle")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```


Lastly, I will calculate and visualize the cumulative area for each mode. I'm almost certainly doing this in an unnecessarily roundabout way, but all the shortcuts I tried resulted in errors.

First, adding up the isochromes for each mode:

```{r}
bike_area <- iso_10min_bike %>%
  mutate(area=st_area(iso_10min_bike)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 
```

```{r}
car_area <- iso_10min_drive %>%
  mutate(area=st_area(iso_10min_drive)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 
```

```{r}
walk_area <- iso_10min_walk %>%
  mutate(area=st_area(iso_10min_walk)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 
```

```{r}
sum(bike_area$bicycle)
sum(car_area$drive)
sum(walk_area$walk)
```
I then created a new data frame from which I could create a bar chart for the three modes:
```{r}
area_totals <- data.frame("mode" = c("Bike", "Car", "Walk"), "Area" = c(55156694, 73836795, 5023677)) 

```

```{r}
ggplot(data=area_totals, aes(x=mode, y=Area)) +
                             geom_bar(stat="identity")
```

